Network Configuration Synthesis with Abstract Topologies

### 2016-11-10 paper/abstraction/paper.tex
pass on 4

\sysname is a system for generating provably-correct templates.  As inputs, it takes a global, network-wide policy (as opposed to a per-device or even per-role policy) and an {\em abstract} topology that describes device roles and
connectivity invariants.  It outputs a template for each role, from which per-device configurations can be generated by instantiating parameters with values drawn from the network topology.
\sysname provides strong guarantees about its templates:
(1) the configurations generated from them correctly implement the specified policy for {\em any} concrete topology that matches the given abstract topology; and (2) if nodes or edges are added or removed in a way compatible
with the abstract topology, only configurations of adjacent devices ever need to be updated.

This section introduces \sysname using the example in Figure~\ref{fig:example}, and the following sections detail template generation and formally define the correctness guarantees.

@@

\subsection{Abstract Compilation}

\sysname builds on the Propane network configuration tool, by supporting both synthesis for and reasoning over parameterized classes of topologies and their interactions with user-specified routing policy.
\sysname takes as input, (1) a topology abstraction in the form of a graph homomorphism annotated with logical constraints about node and edge mutliplicities, and (2) a domain-specific routing policy that consists of a mapping from predicates matching various classes of traffic, to a list of ranked constraints on the paths traffic matching the predicate should prefer to take through the network.

\para{Topology Abstraction}

Consider again the networks from \S\ref{sec:motivation}. Ideally, we would like
to find a suitable topology abstraction that can describe both of the datacenters from figures~\ref{fig:example} and~\ref{fig:example2} as well as any reasonable generalization of these datacenters. \sysname provides several abstraction mechanisms to achieve this goal.

The first is a role-based abstraction that allows an operator to map routers in the concrete network to ``roles" or routers in the abstract network. Figure~\ref{fig:example3} shows an example of an abstraction for both datacenter networks from \S\ref{sec:motivation}. In particular the concrete networks are abstracted into a new topology with 5 different roles: local tor (TL), global tor (TG), local aggregator (AL), global aggregator (AG), and spine (S).

@@

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. Unfortunately this abstraction alone loses a lot of information about the underlying structure of the concrete network. For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of nodes/edges
that may appear in any given concrete network. Operators can then capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ declare that, within any pod P, the number of outgoing edges from $TG$ to $AG$ is equal to the number of nodes that map to $AG$. Similarly, the number of outgoing edges from $AG$ to $TG$ is equal to the number of nodes that map to $TG$. These constraints capture the fact that, within any given pod, the global aggregators and tors are in a full mesh. Furthermore, the constraints $TG = AG$ and $TG \geq 2$ ensure that the number of concrete nodes in role $TG$ is always equal to the number of concrete nodes in role $AG$, which is at least 2. Similar constraints also appear for the local aggregators and tors.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Again, similar constraints appear for the local aggregator role. Finally, the constraints $S \geq 2$, $(S \mod AG) = 0$, and $(S \mod AL) = 0$ ensure that there are at least 2 routers in the spine role, and that the number of spine routers is always divisible by the number of global and local aggregators.
%
The last two constraints are interesting in that they do not involve a simple inequality. In general operators can write constraints using logical formula for any theory supported by modern SMT solvers.

@@

Next let us see how to configure each of the datacenter networks from \S\ref{sec:motivation} by writing a \sysname routing policy over the new abstract topology. In \sysname, we can capture the basic routing behavior for the datacenters with the following constraints:
%
\begin{code}
\Define Routing =
    \$GP  \Path \End(TG)
    \$LP  \Path \End(TL)
    \True \Path \Exit(Peer1 \Prefer Peer2)
\end{code}
\noindent%

@@

\subsection{Abstract Analysis}

Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to. To this end, we develop a sound analysis over abstract topologies to answer questions about the number of disjoint paths between pairs of nodes.

Suppose we want to find a worst-case lower bound on the fault tolerance between any global tor and spine router from figure~\ref{fig:example3} for any concrete topology after applying the routing policy. This can be used as a lightweight verification tool to detect certain classes of configuration bugs. For example, since the network is performing prefix summarization at the border of the datacenter, we might introduce a black hole after enough failures occur to disconnect a tor from a spine router.

We develop a disjoint path analysis works in the following way. For each abstract role in the network, we learn invariants about the number of disjoint paths to routers in that role. These invariants contain a sequence of labels followed by a pair of numbers $(i,j)$. For example, the analysis would start with a fact of the form: $S_P S_{TG} (1,1)$. This means that, there is some group of nodes of size 1 in the TG role in some pod P such that any initial node in the TG role has 1 disjoint path to it (and all such paths are edge-disjoint). Since we know that the each node in the TG role is in a full mesh with nodes in the AG role within any given pod, we might infer a fact for AG of the form: $S_P A_{AG}(e1,1)$. This says that any group of nodes of size e1 in the AG role is reachable with at least 1 edge-disjoint path to each such node from any initial node in the TG role.
%
If we run the abstract analysis over the abstract network from figure~\ref{fig:example3} the analysis will learn a fact of the form $A_{S}(1,2)$, which means that any single spine node is reachable via 2 disjoint paths from any initial global tor node \ryan{it will actually be 1, we need to add a constraint on e1}.

Suppose now that we want to disallow paths with valleys (\IE, paths that go down and then up at some point). Such a constraint can be added to the routing policy in \sysname:
\begin{code}
\Define ValleyFree =
    \True \Path \Novalley(\{TL,TG\},\{AL,AG\},\{S\})
\end{code}
\noindent%

This adds a constraint that applies to all traffic and prevents valleys by defining each level in the datacenter.
Because the abstract analysis is policy sensitive, if we were to run it again with the same abstract topology but the new routing policy, it would only be able to infer that there is a single disjoint path to any given spine router in the worst case. In fact this is exactly the case for the concrete network from figure~\ref{fig:example2}.

>>>>

\sysname is a system for generating provably-correct templates from the network's policy and an {\em abstract} topology that describes device roles and connectivity invariants.
%As inputs, it takes a global, network-wide policy (as opposed to a per-device or even per-role policy) and an {\em abstract} topology that describes device roles and connectivity invariants.  It outputs a template for each role, from which per-device configurations can be generated by instantiating parameters with values drawn from the network topology.
It guarantees that:
(1) the configurations generated from its templates correctly implement the network's policy for {\em any} concrete topology that matches the abstract topology; and (2) if nodes or edges are added or removed in a way compatible
with the abstract topology, only configurations of adjacent devices ever need to be updated.

This section introduces \sysname using the examples in \S\ref{sec:motivation}, and the following sections detail template generation and formally define the correctness guarantees.

@@

\subsection{Inputs}

\sysname takes as input the network's abstract topology and its policy. The policy has two components: the routing policy describes how traffic should flow, and the fault-tolerance policy describes how many simultaneous failures a the network can withstand without loosing connectivity.

\para{Abstract topology}
Abstract topologies in \sysname capture the basic connectivity structure of the network, which holds when the network is originally designed and as it evolves. They are encoded in the form of a graph homomorphism annotated with logical constraints about node and edge multiplicities. We designed these abstractions to be able to precisely capture most real network topologies, while being amenable to the fault tolerance analysis and configuration synthesis that is done by our compiler.

%Ideally, we would like to find a suitable topology abstraction that can describe both of the datacenters from figures~\ref{fig:example} and~\ref{fig:example2} as well as any reasonable generalization of these datacenters. \sysname provides several abstraction mechanisms to achieve this goal.

We introduce our topology abstractions using the networks from \S\ref{sec:motivation}.
The first is a role-based abstraction that allows an operator to map routers in the concrete network to roles in the abstract network. Figure~\ref{fig:example3} shows an example of an abstraction for both datacenter networks from \S\ref{sec:motivation}. In particular the concrete networks are abstracted into a new topology with 5 different roles: local ToR (TL), global ToR (TG), local aggregator (AL), global aggregator (AG), and spine (S).

@@

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. Unfortunately, by itself, this abstraction loses a lot of information about the underlying structure of the concrete network, which would hinder a useful analysis of its fault-tolerance. For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of nodes/edges
that may appear in any given concrete network. Operators can then capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ declare that, within any pod P, the number of outgoing edges from $TG$ to $AG$ is equal to the number of nodes that map to $AG$. Similarly, the number of outgoing edges from $AG$ to $TG$ is equal to the number of nodes that map to $TG$. These constraints capture the fact that, within any given pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $TG = AG$ and $TG \geq 2$ ensure that the number of concrete nodes in role $TG$ is always equal to the number of concrete nodes in role $AG$, which is at least 2. Similar constraints also appear for the local aggregators and ToRs.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Again, similar constraints appear for the local aggregator role. Finally, the constraints $S \geq 2$, $(S \mod AG) = 0$, and $(S \mod AL) = 0$ ensure that there are at least 2 routers in the spine role, and that the number of spine routers is always divisible by the number of global and local aggregators.
%
The last two constraints are interesting in that they do not involve a simple inequality. In general operators can write constraints using logical formula for any theory supported by modern SMT solvers.

@@

\para{Routing policy} Routing policies in \sysname consist of mappings from predicates matching various classes of traffic to a list of ranked constraints on the paths that the traffic matching the predicate should prefer to take through the network. We borrow syntax from Propane~\cite{propane} but, instead of concrete prefixes and nodes, \sysname specifications use abstract prefixes and nodes.

Let us see how to configure each of the datacenter networks from \S\ref{sec:motivation} by writing a \sysname routing policy over the new abstract topology. We can capture the basic routing behavior for the datacenters with the following constraints:
%
\begin{code}
\Define Routing =
    \$GP  \Path \End(TG)
    \$LP  \Path \End(TL)
    \True \Path \Exit(Peer1 \Prefer Peer2)
\end{code}
\noindent%

@@

\para{Fault-tolerance policy} This policy specifies how many link failures the network should be able to withstand before connectivity loss for traffic occurs. Operators can  specify different tolerance levels for different pairs of abstract nodes. For instance, they may say that ToR to spine connectivity should be robust to 2 failures, i.e., no ToR-spine pair should lose connectivity as long as the number of simultaneous link failures is 2 or fewer (for any combination of links); and ToR-to-ToR connectivity should be be robust to 1 failure.

\subsection{Compilation}

%\sysname builds on the Propane network configuration tool, by supporting both synthesis for and reasoning over parameterized classes of topologies and their interactions with user-specified routing policy.

%Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to. 
The \sysname compiler generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on the Propane compiler, which was designed for concrete topologies and policy; we show that it can be extended to abstract inputs and that the extensions are correct.

Second, \sysname evaluates if the fault-tolerance policy can be met, considering the joint impact of the abstract topology and routing policy. Joint analysis is needed, and topology alone is not enough, because connectivity for traffic will suffer even when functional paths exists in the topology if those paths are not allowed by routing policy. It implies that we cannot use standard graph theoretic algorithms here.

Instead, we develop a sound analysis that is based on computing the minimum number of disjoint policy-compliant paths between pairs of nodes over abstract topologies. If this number for a pair of nodes is less than the desired fault-tolerance level, we declare that the policy cannot be met.

Suppose we want to compute the number of of disjoint paths between any global ToR and spine router in Figure~\ref{fig:example3} for any concrete topology after applying the routing policy.
%This can be used as a lightweight verification tool to detect certain classes of configuration bugs. For example, since the network is performing prefix summarization at the border of the datacenter, we might introduce a black hole after enough failures occur to disconnect a tor from a spine router.
We proceed as follows. For each abstract role in the network, we learn invariants about the number of disjoint paths to routers in that role. These invariants contain a sequence of labels followed by a pair of numbers $(i,j)$. For example, the analysis would start with a fact of the form: $S_P S_{TG} (1,1)$. This means that, there is some group of nodes of size 1 in the TG role in some pod P such that any initial node in the TG role has 1 disjoint path to it (and all such paths are edge-disjoint). Since we know that the each node in the TG role is in a full mesh with nodes in the AG role within any given pod, we might infer a fact for AG of the form: $S_P A_{AG}(e1,1)$. This says that any group of nodes of size e1 in the AG role is reachable with at least 1 edge-disjoint path to each such node from any initial node in the TG role.
%
If we run the abstract analysis over the abstract network from Figure~\ref{fig:example3} the analysis will learn a fact of the form $A_{S}(1,2)$, which means that any single spine node is reachable via 2 disjoint paths from any initial global tor node \ryan{it will actually be 1, we need to add a constraint on e1}.

Suppose now that the routing policy disallowed paths with valleys (\IE, paths that go down and then up at some point). 
%Such a constraint can be added to the routing policy in \sysname:
%\begin{code}
%\Define ValleyFree =
%    \True \Path \Novalley(\{TL,TG\},\{AL,AG\},\{S\})
%\end{code}
%\noindent%
%This adds a constraint that applies to all traffic and prevents valleys by defining each level in the datacenter.
Because our analysis is policy sensitive, if we were to run it again with the same abstract topology but the new routing policy, it would only be able to infer that there is a single disjoint path to any given spine router in the worst case. 
In fact this is exactly the case for the concrete network from Figure~\ref{fig:example2}.

Finally, if \sysname deems that the fault-tolerance policy can be met, it generates templates for the abstract topology, which can be instantiated into BGP configurations based on the concrete topology. During concretization, we confirm that the concrete topology adheres to the original abstract topology.
